Association between SARS-CoV-2 and metagenomic content of samples from the Huanan Seafood Market

### 2023-07-13 paper/final_submission/paper.tex
move references from BBL to TEX file

\bibliography{references}
}

>>>>

\begin{thebibliography}{44}

\bibitem[{Bloom}(2021)]{bloom2021recovery}
{Bloom, J.~D.}, 2021 Recovery of deleted deep sequencing data sheds more light
  on the early {Wuhan SARS-CoV-2} epidemic. Molecular Biology and Evolution
  {\bf 38}: 5211--5224.

\bibitem[{Chan {\em et~al.\/}}(2020){Chan, Yuan, Kok, To, Chu, Yang, Xing, Liu,
  Yip, Poon, {\em et~al.\/}}]{chan2020familial}
{Chan, J. F.-W., S.~Yuan, K.-H. Kok, K.~K.-W. To, H.~Chu, {\em et~al.\/}}, 2020
  A familial cluster of pneumonia associated with the 2019 novel coronavirus
  indicating person-to-person transmission: a study of a family cluster. The
  Lancet {\bf 395}: 514--523,
  \url{https://www.thelancet.com/journals/lancet/article/PIIS0140-6736(20)30154-9/fulltext}.

\bibitem[{Chen {\em et~al.\/}}(2020){Chen, Zhou, Dong, Qu, Gong, Han, Qiu,
  Wang, Liu, Wei, {\em et~al.\/}}]{chen2020epidemiological}
{Chen, N., M.~Zhou, X.~Dong, J.~Qu, F.~Gong, {\em et~al.\/}}, 2020
  Epidemiological and clinical characteristics of 99 cases of 2019 novel
  coronavirus pneumonia in {Wuhan, China}: a descriptive study. Lancet {\bf
  395}: 507--513,
  \url{https://www.thelancet.com/journals/lancet/article/PIIS0140-6736(20)30211-7/fulltext}.

\bibitem[{Chen {\em et~al.\/}}(2018){Chen, Zhou, Chen, {\rm and}
  Gu}]{chen2018fastp}
{Chen, S., Y.~Zhou, Y.~Chen, {\rm and} J.~Gu}, 2018 fastp: an ultra-fast
  all-in-one {FASTQ} preprocessor. Bioinformatics {\bf 34}: i884--i890,
  \url{https://academic.oup.com/bioinformatics/article/34/17/i884/5093234}.

\bibitem[{Cohen}(2020)]{cohen2020wuhan}
{Cohen, J.}, 2020 Wuhan seafood market may not be source of novel virus
  spreading globally. Science {\bf 10}: 10.1126/science.abb0611,
  \url{https://www.science.org/content/article/wuhan-seafood-market-may-not-be-source-novel-virus-spreading-globally}.

\bibitem[{Cohen}(2022{a})]{cohen2022anywhere}
{Cohen, J.}, 2022{a} {Anywhere but here: China now insists the pandemic
  didn’t start within its borders. Its scientists are publishing a flurry of
  papers pointing the finger elsewhere}. Science
  \url{https://www.science.org/content/article/pandemic-start-anywhere-but-here-argue-papers-chinese-scientists-echoing-party-line}.

\bibitem[{Cohen}(2022{b})]{cohen2022studies}
{Cohen, J.}, 2022{b} Studies bolster pandemic origin in {W}uhan animal market.
  Science {\bf 375}: 946 -- 947,
  \url{https://www.science.org/doi/epdf/10.1126/science.adb1760}; see Tweet at
  \url{https://twitter.com/sciencecohen/status/1498658469775298568} for image
  of figure from 2022 Chinese CDC pre-print.

\bibitem[{Cohen}(2023{a})]{cohen2023new}
{Cohen, J.}, 2023{a} {A new pandemic origin report is stirring controversy.
  Here are key takeaways}. Science,
  \url{https://www.science.org/content/article/covid-origin-report-controversy}.

\bibitem[{Cohen}(2023{b})]{cohen2023unearthed}
{Cohen, J.}, 2023{b} {Unearthed genetic sequences from China market may point
  to animal origin of COVID-19}. Science
  \url{https://www.science.org/content/article/covid-19-origins-missing-sequences}.

\bibitem[{Courtier-Orgogozo {\rm and} de~Ribera}(2022)]{courtier2022sars}
{Courtier-Orgogozo, V. {\rm and} F.~A. de~Ribera}, 2022 {SARS-CoV-2 infection
  at the Huanan seafood market}. Environmental Research {\bf 214}: 113702,
  \url{https://www.sciencedirect.com/science/article/pii/S0013935122010295}.

\bibitem[{Crits-Christoph {\em et~al.\/}}(2023){Crits-Christoph, Gangavarapu,
  Pekar, Moshiri, Singh, Levy, Goldstein, Suchard, Popescu, Robertson, {\em
  et~al.\/}}]{crits2023genetic}
{Crits-Christoph, A., K.~Gangavarapu, J.~E. Pekar, N.~Moshiri, R.~Singh, {\em
  et~al.\/}}, 2023 {Genetic evidence of susceptible wildlife in SARS-CoV-2
  positive samples at the Huanan Wholesale Seafood Market, Wuhan: Analysis and
  interpretation of data released by the Chinese Center for Disease Control}.
  Zenodo {\bf DOI 10.5281/zenodo.7754299},
  \url{https://zenodo.org/record/7754299#.ZEUz_uzMLX0}.

\bibitem[{GISAID}(2023)]{gisaid_statement}
{GISAID}, 2023 {GISAID comments on the speculations surrounding data
  availability}.
  \url{https://gisaid.org/statements-clarifications/data-availability/}.

\bibitem[{Grabherr {\em et~al.\/}}(2011){Grabherr, Haas, Yassour, Levin,
  Thompson, Amit, Adiconis, Fan, Raychowdhury, Zeng, {\em
  et~al.\/}}]{grabherr2011trinity}
{Grabherr, M.~G., B.~J. Haas, M.~Yassour, J.~Z. Levin, D.~A. Thompson, {\em
  et~al.\/}}, 2011 Trinity: reconstructing a full-length transcriptome without
  a genome from {RNA-Seq} data. Nature Biotechnology {\bf 29}: 644,
  \url{https://www.nature.com/articles/nbt.1883}.

\bibitem[{He {\rm and} Dunn}(2020)]{he2020evaluating}
{He, M. {\rm and} L.~F. Dunn}, 2020 Evaluating incidence and impact estimates
  of the coronavirus outbreak from official and non-official chinese data
  sources. SSRN {\bf DOI 10.2139/ssrn.3540636},
  \url{https://papers.ssrn.com/sol3/Papers.cfm?abstract_id=3540636}.

\bibitem[{Huang {\em et~al.\/}}(2020){Huang, Wang, Li, Ren, Zhao, Hu, Zhang,
  Fan, Xu, Gu, {\em et~al.\/}}]{huang2020clinical}
{Huang, C., Y.~Wang, X.~Li, L.~Ren, J.~Zhao, {\em et~al.\/}}, 2020 Clinical
  features of patients infected with 2019 novel coronavirus in {Wuhan, China}.
  Lancet {\bf 395}: 497--506,
  \url{https://www.thelancet.com/article/S0140-6736(20)30183-5/fulltext}.

\bibitem[{Hvistendahl {\rm and} Mueller}(2023)]{hvistendahl2023}
{Hvistendahl, M. {\rm and} B.~Mueller}, 2023 Chinese censorship is quietly
  rewriting the {Covid-19} story. The New York Times
  \url{https://www.nytimes.com/2023/04/23/world/europe/chinese-censorship-covid.html}.

\bibitem[{Kang {\em et~al.\/}}(2020){Kang, Cheng, {\rm and}
  McNeil}]{kang2020china}
{Kang, D., M.~Cheng, {\rm and} S.~McNeil}, 2020 China clamps down in hidden
  hunt for coronavirus origins.
  \url{https://apnews.com/article/united-nations-coronavirus-pandemic-china-only-on-ap-bats-24fbadc58cee3a40bca2ddf7a14d2955},
  The actual China State Council order described in the article is at
  \url{https://www.documentcloud.org/documents/7340337-State-Research-regulations.html}.

\bibitem[{Li}(2018)]{li2018minimap2}
{Li, H.}, 2018 Minimap2: pairwise alignment for nucleotide sequences.
  Bioinformatics {\bf 34}: 3094--3100,
  \url{https://academic.oup.com/bioinformatics/article/34/18/3094/4994778}.

\bibitem[{Li {\em et~al.\/}}(2020){Li, Guan, Wu, Wang, Zhou, Tong, Ren, Leung,
  Lau, Wong, {\em et~al.\/}}]{li2020early}
{Li, Q., X.~Guan, P.~Wu, X.~Wang, L.~Zhou, {\em et~al.\/}}, 2020 {Early
  transmission dynamics in Wuhan, China, of novel coronavirus--infected
  pneumonia}. New England Journal of Medicine {\bf 382}: 1199--1207,
  \url{https://www.nejm.org/doi/full/10.1056/nejmoa2001316}.

\bibitem[{Liu {\em et~al.\/}}(2022){Liu, Liu, Lei, Jia, He, Liu, Shi, Tan, Zou,
  {\em et~al.\/}}]{liu2022surveillance}
{Liu, W., P.~Liu, W.~Lei, Z.~Jia, X.~He, {\em et~al.\/}}, 2022 {Surveillance of
  SARS-CoV-2 in the environment and animal samples of the Huanan Seafood
  Market}. Research Square {\bf DOI 10.21203/rs.3.rs-1370392/v1},
  \url{https://www.researchsquare.com/article/rs-1370392/v1}.

\bibitem[{Liu {\em et~al.\/}}(2023{a}){Liu, Liu, Lei, Jia, He, Shi, Tan, Zou,
  Wong, Wang, {\em et~al.\/}}]{liu2023surveillance}
{Liu, W.~J., P.~Liu, W.~Lei, Z.~Jia, X.~He, {\em et~al.\/}}, 2023{a}
  {Surveillance of SARS-CoV-2 at the Huanan Seafood Market}. Nature {\bf DOI:
  10.1038/s41586-023-06043-2},
  \url{https://www.nature.com/articles/s41586-023-06043-2}.

\bibitem[{Liu {\em et~al.\/}}(2023{b}){Liu, Liu, Lei, Jia, He, Shi, Tan, Zou,
  Wong, Wang, {\em et~al.\/}}]{liu2023surveillance-chinaxiv}
{Liu, W.~J., P.~Liu, W.~Lei, Z.~Jia, X.~He, {\em et~al.\/}}, 2023{b}
  {Surveillance of SARS-CoV-2 at the Huanan Seafood Market}. ChinaXiv {\bf DOI:
  10.12074/202303.10351V1}, \url{https://chinaxiv.org/abs/202303.10351V1}.

\bibitem[{Ma}(2020)]{ma2020SCMP}
{Ma, J.}, 2020 {Coronavirus: China’s first confirmed Covid-19 case traced
  back to November 17}. South China Morning Post,
  \url{https://www.scmp.com/news/china/society/article/3074991/coronavirus-chinas-first-confirmed-covid-19-case-traced-back},
  archived at
  \url{https://web.archive.org/web/20200315011702/https://www.scmp.com/news/china/society/article/3074991/coronavirus-chinas-first-confirmed-covid-19-case-traced-back}.

\bibitem[{M{\"o}lder {\em et~al.\/}}(2021){M{\"o}lder, Jablonski, Letcher,
  Hall, Tomkins-Tinch, Sochat, Forster, Lee, Twardziok, Kanitz, {\em
  et~al.\/}}]{molder2021sustainable}
{M{\"o}lder, F., K.~P. Jablonski, B.~Letcher, M.~B. Hall, C.~H. Tomkins-Tinch,
  {\em et~al.\/}}, 2021 Sustainable data analysis with {Snakemake}.
  F1000Research {\bf 10}, \url{https://f1000research.com/articles/10-33}.

\bibitem[{Mueller}(2023)]{mueller2023nytimes}
{Mueller, B.}, 2023 New data links pandemic’s origins to raccoon dogs at
  {Wuhan} market. The New York Times
  \url{https://www.nytimes.com/2023/03/16/science/covid-wuhan-market-raccoon-dogs-lab-leak.html}.

\bibitem[{Nishiura {\em et~al.\/}}(2020){Nishiura, Linton, {\rm and}
  Akhmetzhanov}]{nishiura2020initial}
{Nishiura, H., N.~M. Linton, {\rm and} A.~R. Akhmetzhanov}, 2020 {Initial
  cluster of novel coronavirus (2019-nCoV) infections in Wuhan, China is
  consistent with substantial human-to-human transmission}. Journal of Clinical
  Medicine {\bf 9}: 488, \url{https://www.mdpi.com/2077-0383/9/2/488}.

\bibitem[{{ODNI}}(2022)]{odni2022}
{{ODNI}}, 2022 Unclassified summary of assessment on {COVID-19} origins.
  \url{https://www.dni.gov/files/ODNI/documents/assessments/Unclassified-Summary-of-Assessment-on-COVID-19-Origins.pdf}.

\bibitem[{Ondov {\em et~al.\/}}(2016){Ondov, Treangen, Melsted, Mallonee,
  Bergman, Koren, {\rm and} Phillippy}]{ondov2016mash}
{Ondov, B.~D., T.~J. Treangen, P.~Melsted, A.~B. Mallonee, N.~H. Bergman, {\em
  et~al.\/}}, 2016 Mash: fast genome and metagenome distance estimation using
  {MinHash}. Genome biology {\bf 17}: 1--14,
  \url{https://genomebiology.biomedcentral.com/articles/10.1186/s13059-016-0997-x}.

\bibitem[{Pekar {\em et~al.\/}}(2021){Pekar, Worobey, Moshiri, Scheffler, {\rm
  and} Wertheim}]{pekar2021timing}
{Pekar, J., M.~Worobey, N.~Moshiri, K.~Scheffler, {\rm and} J.~O. Wertheim},
  2021 Timing the {SARS-CoV-2} index case in hubei province. Science {\bf 372}:
  412--417, \url{https://www.science.org/doi/full/10.1126/science.abf8003}.

\bibitem[{Pekar {\em et~al.\/}}(2022){Pekar, Magee, Parker, Moshiri,
  Izhikevich, Havens, Gangavarapu, Malpica~Serrano, Crits-Christoph, Matteson,
  {\em et~al.\/}}]{pekar2022molecular}
{Pekar, J.~E., A.~Magee, E.~Parker, N.~Moshiri, K.~Izhikevich, {\em et~al.\/}},
  2022 The molecular epidemiology of multiple zoonotic origins of {SARS-CoV-2}.
  Science {\bf 377}: 960--966,
  \url{https://www.science.org/doi/full/10.1126/science.abp8337}.

\bibitem[{Phan {\em et~al.\/}}(2020){Phan, Nguyen, Luong, Nguyen, Nguyen, Le,
  Nguyen, Cao, {\rm and} Pham}]{phan2020importation}
{Phan, L.~T., T.~V. Nguyen, Q.~C. Luong, T.~V. Nguyen, H.~T. Nguyen, {\em
  et~al.\/}}, 2020 Importation and human-to-human transmission of a novel
  coronavirus in vietnam. New England Journal of Medicine {\bf 382}: 872--874,
  \url{https://www.nejm.org/doi/full/10.1056/nejmc2001272}.

\bibitem[{Pipes {\em et~al.\/}}(2021){Pipes, Wang, Huelsenbeck, {\rm and}
  Nielsen}]{pipes2021assessing}
{Pipes, L., H.~Wang, J.~P. Huelsenbeck, {\rm and} R.~Nielsen}, 2021 Assessing
  uncertainty in the rooting of the {SARS-CoV-2} phylogeny. Molecular Biology
  and Evolution {\bf 38}: 1537--1543,
  \url{https://academic.oup.com/mbe/article/38/4/1537/6028993}.

\bibitem[{Plaue}(2020)]{theilsen}
{Plaue, M.}, 2020 {A correlation measure based on Theil-Sen regression}.
  Torwards Data Science,
  \url{https://towardsdatascience.com/a-correlation-measure-based-on-theil-sen-regression-31b8b9ed64f1}.

\bibitem[{ProMED}(2019)]{ProMED2109}
{ProMED}, 2019 Undiagnosed pneumonia---{China (Hubei)}: Request for
  information. \url{https://promedmail.org/promed-post/?id=6864153}.

\bibitem[{van Dorp {\em et~al.\/}}(2020){van Dorp, Acman, Richard, Shaw, Ford,
  Ormond, Owen, Pang, Tan, Boshier, {\em et~al.\/}}]{van2020emergence}
{van Dorp, L., M.~Acman, D.~Richard, L.~P. Shaw, C.~E. Ford, {\em et~al.\/}},
  2020 Emergence of genomic diversity and recurrent mutations in {SARS-CoV-2}.
  Infection, Genetics and Evolution {\bf 83}: 104351,
  \url{https://www.sciencedirect.com/science/article/pii/S1567134820301829}.

\bibitem[{VanderPlas {\em et~al.\/}}(2018){VanderPlas, Granger, Heer, Moritz,
  Wongsuphasawat, Satyanarayan, Lees, Timofeev, Welsh, {\rm and}
  Sievert}]{vanderplas2018altair}
{VanderPlas, J., B.~Granger, J.~Heer, D.~Moritz, K.~Wongsuphasawat, {\em
  et~al.\/}}, 2018 Altair: interactive statistical visualizations for python.
  Journal of open source software {\bf 3}: 1057,
  \url{https://joss.theoj.org/papers/10.21105/joss.01057}.

\bibitem[{{WHO}}(2020)]{who2020}
{{WHO}}, 2020 {Pneumonia of unknown cause -- China}.
  \url{https://www.who.int/emergencies/disease-outbreak-news/item/2020-DON229}.

\bibitem[{Wood {\em et~al.\/}}(2019){Wood, Lu, {\rm and}
  Langmead}]{wood2019improved}
{Wood, D.~E., J.~Lu, {\rm and} B.~Langmead}, 2019 Improved metagenomic analysis
  with {K}raken 2. Genome Biology {\bf 20}: 1--13.

\bibitem[{Worobey {\em et~al.\/}}(2022){Worobey, Levy, Malpica~Serrano,
  Crits-Christoph, Pekar, Goldstein, Rasmussen, Kraemer, Newman, Koopmans, {\em
  et~al.\/}}]{worobey2022huanan}
{Worobey, M., J.~I. Levy, L.~Malpica~Serrano, A.~Crits-Christoph, J.~E. Pekar,
  {\em et~al.\/}}, 2022 {The Huanan Seafood Wholesale Market in Wuhan was the
  early epicenter of the COVID-19 pandemic}. Science {\bf 377}: 951--959,
  \url{https://www.science.org/doi/10.1126/science.abp8715}.

\bibitem[{Wu}(2023)]{wu2023atlantic}
{Wu, K.~J.}, 2023 The strongest evidence that an animal started the pandemic.
  The Atlantic
  \url{https://www.theatlantic.com/science/archive/2023/03/covid-origins-research-raccoon-dogs-wuhan-market-lab-leak/673390/}.

\bibitem[{{Wuhan Municipal Health Commission}}(2019)]{healthcommission2019}
{{Wuhan Municipal Health Commission}}, 2019 {Wuhan Municipal Health
  Commission's briefing on the current situation of pneumonia in our city}. The
  original report has been deleted, but archived versions are available on the
  Wayback Machine at
  \url{https://web.archive.org/web/20200131202951/http:/wjw.wuhan.gov.cn/front/web/showDetail/2019123108989}.

\bibitem[{{Wuhan Municipal Health Commission}}(2020)]{healthcommission2020}
{{Wuhan Municipal Health Commission}}, 2020 {Wuhan Municipal Health
  Commission's report on viral pneumonia of unknown cause}. The original report
  has been deleted, but archived versions are available on the Wayback Machine
  at
  \url{https://web.archive.org/web/20200111023106/http://wjw.wuhan.gov.cn/front/web/showDetail/2020011109035}.

\bibitem[{Xiao {\em et~al.\/}}(2021){Xiao, Newman, Buesching, Macdonald, {\rm
  and} Zhou}]{xiao2021animal}
{Xiao, X., C.~Newman, C.~D. Buesching, D.~W. Macdonald, {\rm and} Z.-M. Zhou},
  2021 {Animal sales from Wuhan wet markets immediately prior to the COVID-19
  pandemic}. Scientific Reports {\bf 11}: 1--7,
  \url{https://www.nature.com/articles/s41598-021-91470-2}.

\bibitem[{Zhang {\em et~al.\/}}(2020){Zhang, Tan, Ling, Lu, Liu, Yi, Jia, Wu,
  Shi, Xu, {\em et~al.\/}}]{zhang2020viral}
{Zhang, X., Y.~Tan, Y.~Ling, G.~Lu, F.~Liu, {\em et~al.\/}}, 2020 Viral and
  host factors related to the clinical outcome of {COVID-19}. Nature {\bf 583}:
  437--440, \url{https://www.nature.com/articles/s41586-020-2355-0}.

\end{thebibliography}
